# Peak Calling {#calling}

> Peak caller: [MACS2](https://github.com/macs3-project/MACS)

## No. of peaks
```{r calling-stat}
knitr:: kable(dba.show(dbo),caption='Peak Calling Statistic')
```

```{r calling-stat-fig1, fig.cap="No. of peaks"}
ggplot(df.dbo, aes(Factor, Intervals)) + 
geom_boxplot(width=.3,outlier.shape=NA,alpha=.7) + 
geom_jitter(width=.1, size=4,alpha=.5) +
ylab("No. of peaks") + theme_bw()
```

```{r calling-stat-fig2, fig.cap="No. of peaks (per million mapped read)"}
df.dbo.mapping<-merge(df.dbo, dt.mapping, by.x="ID", by.y="SampleName")
ggplot(df.dbo.mapping, aes(Factor, Intervals/`uniquely mapped`*1e+6)) + 
geom_boxplot(width=.3,outlier.shape=NA,alpha=.7) + 
geom_jitter(width=.1, size=4,alpha=.5) + 
ggsignif::geom_signif(comparisons = list(c("FD","FV"),c("MD","MV")),size=.9,textsize=5.5) +
ylab("No. of peaks\n(per million mapped read)") + theme_bw()
```

```{r calling-stat-fig3, fig.cap="No. of peak in the promoter regions"}
if(!file.exists("RData/dt.peak.num.csv")){
    gl.peak[["sample"]]<-lapply(dbo$peaks,makeGRangesFromDataFrame,keep.extra.columns=T)
    names(gl.peak[["sample"]])<-dba.show(dbo)$ID
    sapply(gl.peak[['sample']],length)
    li.peakAnno[['sample']]<-lapply(gl.peak[["sample"]], ChIPseeker::annotatePeak, tssRegion=c(-2e3, 2e3), TxDb=txdb, annoDb="org.Hs.eg.db", verbose=F) # list of 'csAnno'
    li.peakAnno[['sample']]
    dl.peakAnno[['sample']]<-lapply(li.peakAnno[['sample']], as.data.table)
    li.peak.chr=lapply(li.peakAnno[['sample']], function(i) as.data.table(i)[,.N,seqnames])
    li.peak.promo=lapply(li.peakAnno[['sample']], function(i) as.data.table(i)[grepl("Promoter",annotation),.N,seqnames])

    dt.peak.num<-data.table(
        `ID`=dba.show(dbo)$ID,
        `TSS_peak`=sapply(li.peak.promo,function(i) i[,sum(N)]), # No. of TSS peak
        `TSS_chrX_peak`=sapply(li.peak.promo,function(i) i[seqnames=="X",sum(N)]), # No. of chrX TSS peak
        `TSS_autosomal_peak`=sapply(li.peak.promo,function(i) i[!seqnames%in%c("X","Y"),sum(N)]), # No. of autosomal TSS peak
        `chrX_peak`=sapply(li.peak.chr,function(i) i[seqnames=="X",sum(N)]), # No. of chrX peak
        `autosomal_peak`=sapply(li.peak.chr,function(i) i[!seqnames%in%c("X","Y"),sum(N)]) # No. of autosomal peak
        )
    fwrite(dt.peak.num, file="RData/dt.peak.num.csv")
}else{
    dt.peak.num<-fread("RData/dt.peak.num.csv")
}

ggplot(merge(df.dbo.mapping,dt.peak.num), aes(Factor, TSS_peak/`uniquely mapped`*1e+6)) + 
geom_boxplot(width=.3,outlier.shape=NA,alpha=.7) + 
geom_jitter(width=.1, size=4,alpha=.5) +
ggsignif::geom_signif(comparisons = list(c("FD","FV"),c("MD","MV")),size=.9,textsize=5.5) +
ylab("No. of peaks in promoter\n(per million mapped read)") + theme_bw()

if(FALSE){
    ggplot(merge(df.dbo.mapping,dt.peak.num), aes(Factor, TSS_chrX_peak/`uniquely mapped`*1e+6)) + 
        geom_boxplot(width=.3,outlier.shape=NA,alpha=.7) + 
        ylab("No. of peaks in chrX promoter\n(per million read)") +
        geom_jitter(width=.1, size=4,alpha=.5) + theme_bw()

    ggplot(merge(df.dbo.mapping,dt.peak.num), aes(Factor, TSS_autosomal_peak/`uniquely mapped`*1e+6)) + 
        geom_boxplot(width=.3,outlier.shape=NA,alpha=.7) + 
        ylab("No. of peaks in autosomal promoter (per million read)") +
        geom_jitter(width=.1, size=4,alpha=.5) + theme_bw()

    ggplot(merge(df.dbo.mapping,dt.peak.num), aes(Factor, chrX_peak/`uniquely mapped`*1e+6)) + 
        geom_boxplot(width=.3,outlier.shape=NA,alpha=.7) + 
        ylab("No. of peaks in chrX\n(per million read)") +
        geom_jitter(width=.1, size=4,alpha=.5) + theme_bw()

    ggplot(merge(df.dbo.mapping,dt.peak.num), aes(Factor, autosomal_peak/`uniquely mapped`*1e+6)) + 
        geom_boxplot(width=.3,outlier.shape=NA,alpha=.7) + 
        ylab("No. of peaks in autosomes\n(per million read)") +
        geom_jitter(width=.1, size=4,alpha=.5) + theme_bw()
}

```

## No. of peaks by the consensus level 

### All samples (n=`r nrow(df.dbo)`)
```{r peaks-by-consensus-level1, fig.cap="No of peaks by the consensus level"}
print(olap.rate<-dba.overlap(dbo,mode=DBA_OLAP_RATE)) #  No. of peaks by minOverlap (1..)
plot(olap.rate,type='b',ylab='# peaks', xlab='Overlap at least this many peaksets')
```

### FV (n=`r nrow(df.dbo[df.dbo$Factor=="FV",])`)
```{r peaks-by-consensus-level-FV, fig.cap="No of peaks by the consensus level (FV)"}
print(olap.rate<-dba.overlap(dbo,dbo$masks$FV,mode=DBA_OLAP_RATE)) #  No. of peaks by minOverlap (1..)
plot(olap.rate,type='b',ylab='# peaks', xlab='Overlap at least this many peaksets')
```

### FD (n=`r nrow(df.dbo[df.dbo$Factor=="FD",])`)
```{r peaks-by-consensus-level-FD, fig.cap="No of peaks by the consensus level (FD)"}
print(olap.rate<-dba.overlap(dbo,dbo$masks$FD,mode=DBA_OLAP_RATE)) #  No. of peaks by minOverlap (1..)
plot(olap.rate,type='b',ylab='# peaks', xlab='Overlap at least this many peaksets')
```

### MV (n=`r nrow(df.dbo[df.dbo$Factor=="MV",])`)
```{r peaks-by-consensus-level-MV, fig.cap="No of peaks by the consensus level (MV)"}
print(olap.rate<-dba.overlap(dbo,dbo$masks$MV,mode=DBA_OLAP_RATE)) #  No. of peaks by minOverlap (1..)
plot(olap.rate,type='b',ylab='# peaks', xlab='Overlap at least this many peaksets')
```

### MD (n=`r nrow(df.dbo[df.dbo$Factor=="MD",])`)
```{r peaks-by-consensus-level-MD, fig.cap="No of peaks by the consensus level (MD)"}
print(olap.rate<-dba.overlap(dbo,dbo$masks$MD,mode=DBA_OLAP_RATE)) #  No. of peaks by minOverlap (1..)
plot(olap.rate,type='b',ylab='# peaks', xlab='Overlap at least this many peaksets')
```

## Consensus peaks across factors of interests {#consensus_peak}
- A peak should be supported by **at least `r config$minOverlap` samples** within each of the factor group (e.g. FD, FV)
```{r consensus-peak-table1}
DT::datatable(dba.show(dbo.con), caption="No of consensus peaks")

print(olap.rate<-dba.overlap(dbo,dbo$masks$FV,mode=DBA_OLAP_RATE)) #  No. of peaks by minOverlap (1..)
```

- **Final number of merged peaks: `r dbo.cnt$totalMerged`**
- This number of peaks were used for [differential binding region analysis (DBR)](#diffbind)

## Fraction of Read in Peaks (FRiP)
- The proportion of reads that overlap a peak in the consensus peakset.
- It can be used to indicate which samples show more enrichment overall.
```{r frip1}
ggplot(dba.show(dbo.cnt), aes(Factor, FRiP)) + 
geom_boxplot(width=.3,outlier.shape=NA,alpha=.7) + 
geom_jitter(width=.1, size=4,alpha=.5) +
ggsignif::geom_signif(comparisons = list(c("FD","FV"),c("MD","MV")),size=.9,textsize=5.5) +
ylab("Fraction of Read in Peak") + theme_bw()
```

## Heatmap
```{r peak-heatmap1, fig.cap="Heatmap based on the affinity matrix (aka. count-table)"}
dba.plotHeatmap(dbo.cnt,main=paste("Based on",dbo.cnt$totalMerged,"merged peaks"))
```

## PCA
```{r peak-pca1, fig.cap="PCA plot based on the affinity matrix (aka. count-table)"}
dba.plotPCA(dbo.cnt, attributes=DBA_FACTOR, label=DBA_ID) # based on binding affinity matrix
```
