# Differential Binding Region (MD vs MV; n=5 each) {#diffbind-md-vs-mv-n-10}

> Based on `r dbo.cnt$totalMerge` consensus peaks (see [here](#consensus_peak))
>
> Software package used: [DiffBind](https://bioconductor.org/packages/release/bioc/html/DiffBind.html)
>
> Both **M2D** and **M2V** were removed from the analysis

> [`Background normalisation`](https://bioconductor.org/packages/3.12/bioc/vignettes/DiffBind/inst/doc/DiffBind.pdf#page=38) method using `DiffBind3`

```{r diffbind3-bg-norm-m10, echo=T,message=T}
li.diff
li.diff[['bg']][["male-n10"]]<-dba(dbo.cnt,dbo.cnt$masks$M & !dbo.cnt$mask$Replicate.2) # isa DBA

li.diff[['bg']]['male-n10']<-lapply(li.diff[['bg']]['male-n10'], dba.normalize, method=DBA_ALL_METHODS,normalize=DBA_NORM_NATIVE,library=DBA_LIBSIZE_FULL,background=TRUE)
li.diff[['bg']]['male-n10']<-lapply(li.diff[['bg']]['male-n10'], dba.contrast, design="~Replicate+Factor")
li.diff[['bg']]['male-n10']<-lapply(li.diff[['bg']]['male-n10'], dba.analyze, method=DBA_ALL_METHODS)

lapply(li.diff[['bg']]['male-n10'], dba.show, bContrasts=TRUE)
lapply(li.diff[['bg']]['male-n10'], dba.report, method=DBA_ALL_METHODS,bDB=T,bGain=T,bLoss=T)
lapply(li.diff[['bg']]['male-n10'], dba.normalize, bRetrieve=T,method=DBA_ALL_METHODS)

```
## No. of DBR
### adjutedd P-val<0.05
```{r diff-bind-num1-m10}
knitr::kable(dba.show(li.diff[['bg']][['male-n10']],bContrasts=T),caption="Number of differential binding regions")
```

```{r diff-bind-venn1-m10, fig.cap="DBR between edgeR & DESeq2 (MD vs MV)"}
print(li.diff.res<-lapply(li.diff[['bg']]['male-n10'], dba.report,method=DBA_ALL_METHODS,bDB=TRUE,bAll=TRUE)) # isa DBA
dba.plotVenn(li.diff.res[["male-n10"]],1:2, main="",sub="")
```

```{r diff-bind-pca1-m10, fig.cap="PCA based on DBR (MD vs MV)"}
dba.plotPCA(li.diff[['bg']][["male-n10"]], attributes=DBA_FACTOR, label=DBA_ID) 
```

### adjutedd P-val<0.01:
```{r diff-bind-num2-m-m10}
knitr::kable(
             dba.show(li.diff[['bg']][['male-n10']],bContrasts=T,th=.01)
             ,caption="Number of differential binding regions"
)
```

## DBR from DESeq2
```{r dbr-annotation-deseq2-m-m10}
fwrite(data.table(dba.report(li.diff[['bg']][['male-n10']],th=1, method=DBA_DESEQ2, DataType=DBA_DATA_FRAME,bCounts=T))[order(Chr,Start)],file="RData/DBR_analysis_unfiltered_male_n10<-DESeq2.csv.gz")

gl.dbr.DESeq2.n10<-lapply(li.diff[['bg']]['male-n10'], dba.report, method=DBA_DESEQ2, DataType=DBA_DATA_GRANGES)
li.peakAnno[['dbr']][['DESeq2']]['male-n10']<-lapply(gl.dbr.DESeq2.n10, ChIPseeker::annotatePeak, tssRegion=c(-2e3, 2e3), TxDb=txdb, annoDb="org.Hs.eg.db", verbose=F) # list of 'csAnno'
```
- Unfiltered list of DBR: [MD vs MF (n=5 each)](RData/DBR_analysis_unfiltered_male_n10_DESeq2.csv.gz)

### Normalisation factors
```{r dbr-norm-factor-deseq2-m-m10, fig.cap="Normalisation factors by group"}
df.norm<-
    cbind(dba.show(li.diff[['bg']][['male-n10']]),
            `Norm factor`=li.diff[['bg']][['male-n10']][['norm']][['DESeq2']]$norm.facs)

ggplot(df.norm, aes(Condition, `Norm factor`)) + 
geom_boxplot(width=.3,outlier.shape=NA,alpha=.7) + 
geom_jitter(width=.1, size=4,alpha=.5) + 
facet_wrap(~Tissue) +
ggsignif::geom_signif(comparison=list(c("D","V")),size=.9,textsize=5.5) +
ylab("Normalisation factor") + theme_bw()

```

### Male (MD vs MV)
```{r diff-bind-list-deseq2-m-m10}
DT::datatable(
              data.frame(li.peakAnno[['dbr']][['DESeq2']][['male-n10']]@anno)
              ,extensions="Buttons",
              rownames=F,
              colnames=c('Chr'='seqnames'),
              caption="DBR of MD vs MV (DESeq2; FDR<0.05)",
              options=list(pageLength=5)
)
```

```{r plotMA-male-deseq2-fig1-m10, fig.cap="MA plot (DESeq2)"}
dba.plotMA(li.diff[['bg']][["male-n10"]], method=DBA_DESEQ2,th=0.05,bXY=F,bSmooth=F,dotSize=1.5) 
```

```{r plotVol-male-deseq2-fig1-m10, fig.cap="Volcano plot (DESeq2)"}
dba.plotVolcano(li.diff[['bg']][["male-n10"]], method=DBA_DESEQ2, dotSize=1.5,th=0.05)
```

## DBR from edgeR
```{r dbr-annotation1-m10}
fwrite(data.table(dba.report(li.diff[['bg']][['male-n10']],th=1, method=DBA_EDGER, DataType=DBA_DATA_FRAME,bCounts=F))[order(Chr,Start)],file="RData/DBR_analysis_unfiltered_male_n10_edgeR.csv.gz")

gl.dbr.edgeR.n10<-lapply(li.diff[['bg']]['male-n10'], dba.report, method=DBA_EDGER, DataType=DBA_DATA_GRANGES)
li.peakAnno[['dbr']][['edgeR']]['male-n10']<-lapply(gl.dbr.edgeR.n10, ChIPseeker::annotatePeak, tssRegion=c(-2e3, 2e3), TxDb=txdb, annoDb="org.Hs.eg.db", verbose=F) # list of 'csAnno'
```

### Normalisation factors
```{r dbr-norm-factor-edger-m10, fig.cap="Normalisation factors by group"}
df.norm<-cbind(dba.show(li.diff[['bg']][['male']]),
            `Norm factor`=li.diff[['bg']][['male']][['norm']][['edgeR']]$norm.facs)

ggplot(df.norm, aes(Condition, `Norm factor`)) + 
geom_boxplot(width=.3,outlier.shape=NA,alpha=.7) + 
geom_jitter(width=.1, size=4,alpha=.5) + 
facet_wrap(~Tissue) +
ggsignif::geom_signif(comparison=list(c("D","V")),size=.9,textsize=5.5) +
ylab("Normalisation factor") + theme_bw()

```

### Male (MD vs MV)
```{r diff-bind-list-m-m10}
DT::datatable(
              data.frame(li.peakAnno[['dbr']][['edgeR']][['male-n10']]@anno)
              ,extensions="Buttons",
              rownames=F,
              colnames=c('Chr'='seqnames'),
              caption="DBR of MD vs MV (edgeR; FDR<0.05)",
              options=list(pageLength=5)
)
```

```{r plotMA-male-edgeR-fig1-m10, fig.cap="MA plot (edgeR)"}
dba.plotMA(li.diff[['bg']][["male-n10"]], method=DBA_EDGER,th=0.05,bXY=F,bSmooth=F,dotSize=1.5,bNormalized=T)
```

```{r plotVol-male-edgeR-fig1-m10, fig.cap="Volcano plot (edgeR)"}
dba.plotVolcano(li.diff[['bg']][["male-n10"]], method=DBA_EDGER, dotSize=1.5, th=0.05, fold=log2(2))
```

- Unfiltered list of DBR: [MD vs MF (n=5 each)](RData/DBR_analysis_unfiltered_male_n10_edgeR.csv.gz)
